Per capita consumption

### 2022-06-23 PerCapitaConsumption.Rmd
added new totcons but wont knit

---
title: "Per captia consumption"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

@@

## Diet info for multispecies modeling

@@

@@

The function below should return per-capita consumption by age class. This can then be converted to lengthbin as necessary.

@@

## NOBA

@@

## CCA

@@

>>>>

---
title: "Per capita consumption"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

@@

# Consumption info for multispecies modeling

@@

## Approach 1 based on detailedDietCheck.txt

@@

The function below should return per-capita consumption by age class using the `detailedDietCheck.txt` output. This can then be converted to lengthbin as necessary.

@@

### NOBA

@@

### CCA

@@

## Approach 2 based on PROD.nc

An existing `atlantisom` function `calc_pred_diet()` was [tested and found to produce consumption estimates orders of magnitude higher](https://sgaichas.github.io/poseidon-dev/NOBAmsdiets.html) than those derived from `detailedDietCheck.txt`.

Here we develop new functions that don't apply global diet composition to consumption but rather just get total consumption from PROD.nc. The challenge is figuring out what the units are of "Eat" and "Grazing" and how to appropriately expand them to the population level (water volume or numbers in a polygon).

First we need to get "Eat" and "Grazing" from PROD.nc and "Vol" from the .nc file. Here is our model setup from the config file.  Lets look at NOBA_sacc_38, which is being used for mskeyrun simulated dataset generation. **THIS OUTPUT IS NOT COMPARABLE TO NOBA_March_2020 above**.

# WARNING!
# hardcoding because making copies of this in each repo is too cumbersome
# this currently lives in mskeyrun to make the simulated datasets with no climate
# change as appropriate!

dir <- "/Users/sarah.gaichas/Documents/0_Data/ms-keyrun/simulated-data/atlantisoutput/NOBA_sacc_38"

file_fgs <- "nordic_groups_v04.csv"
file_biolprm <- "nordic_biol_incl_harv_v_011_1skg.prm"
file_bgm <- "Nordic02.bgm"
file_init <- "nordic_biol_v23.nc"
file_runprm <- "nordic_run_v01.xml"
scenario <- "nordic_runresults_01"

select_groups <- mskeyrun::simFocalSpecies$Name # 11 mskeyrun fish species

verbose <- FALSE

This code is part of `atlantisom::run_truth()`. With our selected species from `mskeyrun`, only `eat` will be created. `grazing` applies to non-age structured groups.

**If this works for fish, test for inverts too**

  # Read in the functional groups csv since that is used by many functions
  fgs <- load_fgs(dir = dir, file_fgs = file_fgs)
  # Read in the biomass pools
  bps <- load_bps(dir = dir, fgs = file_fgs, file_init = file_init)
  # Read in the biological parameters
  biol <- load_biolprm(dir = dir, file_biolprm = file_biolprm)
  # Read in the run parameters
  runprm <- load_runprm(dir = dir, file_runprm = file_runprm)

  nc_out <- paste0(scenario, ".nc")
  nc_prod <- paste0(scenario, "PROD.nc")

  # Get the boundary boxes
  allboxes <- load_box(dir = dir, file_bgm = file_bgm)
  boxes <- get_boundary(allboxes)

  eat <- load_nc(dir = dir,
                     file_nc = nc_prod,
                     bps = bps,
                     fgs = fgs,
                     select_groups = select_groups,
                     select_variable = "Eat",
                     check_acronyms = TRUE,
                     bboxes = boxes)
  if(verbose) message("Eaten read in.")

  grazing <- load_nc(dir = dir,
                 file_nc = nc_prod,
                 bps = bps,
                 fgs = fgs,
                 select_groups = select_groups,
                 select_variable = "Grazing",
                 check_acronyms = TRUE,
                 bboxes = boxes)
  if(verbose) message("Grazing read in.")

  vol <- load_nc_physics(dir = dir,
                         file_nc = nc_out,
                         physic_variables = "volume",
                         aggregate_layers = FALSE,
                         bboxes = boxes)
  if(verbose) message("Volume read in.")

Now try the new function `atlantisom::calc_pred_cons()` which uses the volume of all water column layers to expand the atoutput (assumed to be mg N per cubic m at that timestep) to bio_eaten in tons.

This is the function code

```{r, code = readLines("https://github.com/r4atlantis/atlantisom/raw/master/R/calc_pred_cons.R"), eval=F}
```

This is the current result

biomass_eaten <- calc_pred_cons(eat = eat, grazing = grazing, vol = vol, biolprm = biol, runprm = runprm)

codtotcons <- biomass_eaten %>% 
  dplyr::filter(species=="North_atl_cod") %>%
  dplyr::group_by(time, agecl) %>% 
  dplyr::summarise(totcons = sum(bio_eaten))

Plotting a cod example total consumption by age class

```{r}
ggplot2::ggplot(codtotcons, aes(x=time, y=totcons)) + geom_line() + facet_wrap(~agecl, scales="free")
